### 2019-04-05 abstract.tex
upd

\documentclass[11pt,a4paper]{article}

\usepackage[margin=1in]{geometry}
\usepackage{url}
\usepackage{hyperref}
\usepackage{natbib}

\title{Neural dialogue act tagging with transformer pre-training (or something like that...)}

\author{
  First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
Affiliation / Address line 3 \\
  {\tt email@domain} \\\and
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\
}

\date{}

\begin{document}
\maketitle

\begin{abstract}
  abstract
\end{abstract}

\section{Introduction}
\begin{itemize}
  \item \cite{austinHowThingsWords2009} -- Speech acts
  \item \cite{coreCodingDialogsDAMSL1997} -- DAMSL dialogue act scheme
  \item \cite{devlinBERTPretrainingDeep2018} -- BERT
    Main questions:
    1. Are pre-trained transformer encoding (BERT) helpful for representation in a dialogue task (speficially, dialogue act tagging)?
    2. Are additional dialogue specific features helpful (discourse markers, disfuency, laughter)?
\end{itemize}

\section{Related Work}
\begin{itemize}
  \item \cite{cerisaraEffectsUsingWord2vec2017}
  \item \cite{kalchbrennerRecurrentConvolutionalNeural2013}
  \item \cite{pragstVectorRepresentationUtterances2018}
  \item \cite{sordoniNeuralNetworkApproach2015}
  \item \cite{stolckeDialogueActModeling2000} - HMM dialogue act tagging
  \item \cite{khanpourDialogueActClassification}
  \item \cite{tranPreservingDistributionalInformation2017}
\end{itemize}

\section{Model}
The proposed model has two components: an utterance encoder, and a sequence model that predicts the dialogue act tag.
For the sequence model, we use a simple RNN.
Conceptually the hidden state of this RNN represents the discourse context which, 
together with the encoded utterance, is used to predict the discourse act tag.

\section{Experiments}
Our aim is to test the effectiveness of different utterance encoding schemas. 
\begin{enumerate}
  \item Average word embedding (word2vec, FastText)
  \item CNN (with/without word2vec initalization, with/without `freezing' the embeddings )
  \item BERT encoded sentences (with/without additional unsupervised pretraining)
  \item Average BERT word embeddings (?)
  \end{enumerate}

  Pre-trained models benefit from learning on large amounts of online data, however it might be the case that for dialogue act tagging additional information only present in natural dialogue data will be useful. Therefore, we evaluate the impact of disfluencies and non-verbal signals, such as laughters on the performance of our models. \textbf{(How?)} For the datasets where no annotation for disfluencies is available, this information can be predicted from recognised speech \citep{hough2017joint,shalyminov2018multi}.

  Furthermore PoS tags

\bibliography{abstract}{}
\bibliographystyle{acl_natbib}

\end{document}