### 2020-12-01 reports/CLASP seminar/abstract.tex
seminar absract

\documentclass[11pt,a4paper]{article}

\title{Tickle me BERT: Using laughter to probe pre-trained encoders for dialogue act recognition}

\date{February 26, 2020}
\author{Bill Noble \& Vladislav Maraev}

\begin{document}
\maketitle
\pagenumbering{gobble}

We investigate the usefulness of BERT for dialogue act recognition.
We are interested in both how useful the pre-training procedure is and 
how well the model adapts to the dialogue domain.
To confirm that the fine-tuned model learns to represent dialogical features, 
we look at how it uses laughter, a phenomenon specific to dialogue.

\end{document}